---
title: "Assessing Overlap and Covariate Balance"
author: "David M. Vock"
date: "PubH 7485/8485"
output: beamer_presentation
theme: "Boadilla"
colortheme: "whale"
fonttheme: "structurebold"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
source("Rexample_covariate_overlap_library.R")
svyQuant_alt <- function (vars, design, q = 0.5) {
  res <- vector()
  for (i in 1:length(vars)) {
    var <- vars[i]
    res[i] <- oldsvyquantile(design$variables[var], design = design, 
      quantiles = q[1], na.rm = TRUE)
  }
  out <- as.vector(res)
  names(out) <- vars
  out
}

environment(svyQuant_alt) <- asNamespace('tableone')
assignInNamespace("svyQuant", svyQuant_alt, ns = "tableone")

data(GerberGreenImai)
imai <- GerberGreenImai

imai <- mutate(imai,
  VOTE96.1F = factor(VOTE96.1, labels = c("No", "Yes")),
  VOTED98F = factor(VOTED98, labels = c("No", "Yes")),
  NEWF = factor(NEW, labels = c("Previous Voter", "New Voter")),
  MAJORPTYF = factor(MAJORPTY, labels = c("Republican", "Democrat")),
  PERSONSF = factor(PERSONS, labels = c("1 Voter", "2+ Voters")),
  PHN.C1F = factor(PHN.C1, labels = c("Not Contacted", "Contacted")))
imai <- set_variable_labels(imai,
	VOTED98F = "Voted in 1998",
  PHN.C1F = "Contacted by phone", 
  PERSONSF = "Voters in household", 
  WARD = "Ward of residence", 
  AGE = "Age (years)", 
  MAJORPTYF = "Party affilation", 
  VOTE96.1F = "Voted in 1996", 
  NEWF = "New voter" 
)

p1 <- glm(PHN.C1 ~ (PERSONS + VOTE96.1 + NEW + MAJORPTY + AGE), 
  data = imai, family = "binomial")

```

## Unadjusted/Unweighted/Unmatched Table 1

- Using tableone package
```{r, echo = TRUE}
vars <- c("VOTED98F", "PERSONSF", "AGE", "VOTE96.1F", 
  "NEWF", "MAJORPTYF",  "WARD")
tabUnmatched <- CreateTableOne(vars = vars, strata = "PHN.C1F",
  data = imai, test = FALSE)
t1 <- print(tabUnmatched, smd = TRUE, showAllLevels = TRUE, varLabels = TRUE)
```

## Unadjusted/Unweighted/Unmatched Table 1
```{r, echo = TRUE, warning = FALSE, size="tiny", results = "asis"}
knitr::kable(t1)
```

## Other Graphical Measures

- Standardized mean difference is one measure of differences between groups
- SMD does not help assess the positivity assumption (covariate overlap)
- Technically need the entire (joint) distribution to be the same between groups not just measure of central tendancy
- Thus, plotting data is of key importance; use cobalt and WeightIt packages

## Create WeightIt object

- Set it up to (eventually) do propensity score weighting but for now will do unadjusted analyses

```{r, echo = TRUE}
W.out <- weightit(PHN.C1 ~ (PERSONSF + VOTE96.1F + NEWF + MAJORPTYF + AGE), data = imai,
                  method = "ps", estimand = "ATE")
```

## Create Plots of Distribution of Key Variables

```{r, echo = TRUE, size="tiny", fig.height=4}
bal.plot(W.out, var.name = "AGE", which = "unadjusted")
```


## Create Plots of Distribution of Key Variables

```{r, echo = TRUE, size="tiny", fig.height=4}
bal.plot(W.out, var.name = "VOTE96.1F", which = "unadjusted")
```


## Weighted Differences in Key Variables Between Groups
```{r, echo = TRUE, warning=FALSE, results = "hide", size="tiny"} 
ps <- predict(p1, type = "response")
imai$weight <- imai$PHN.C1/ps + (1-imai$PHN.C1)/(1-ps)
imaiSvy <- svydesign(ids = ~ 1, data = imai, weights = ~ weight)

tabWeighted <- svyCreateTableOne(vars = vars, strata = "PHN.C1F",
  data = imaiSvy, test = FALSE)
t2 <- print(tabWeighted, smd = TRUE, showAllLevels = TRUE, varLabels = TRUE)
```

## Weighted Differences in Key Variables Between Groups
```{r, echo = TRUE, warning = FALSE, size="tiny", results = "asis"}
knitr::kable(t2)
```

## Create Plots of Distribution of Key Variables

```{r, echo = TRUE, size="tiny", fig.height=4}
bal.plot(W.out, var.name = "AGE")
```

## Create Plots of Distribution of Key Variables

```{r, echo = TRUE, size="tiny", fig.height=4}
bal.plot(W.out, var.name = "VOTE96.1F")
```

## Create WeightIt object

- Are there are other ways of defining weights? YES! Many available as part of weightit
- Many directly target balance of the covariates

```{r, echo = TRUE, cache= TRUE}
W.out <- weightit(PHN.C1 ~ (PERSONSF + VOTE96.1F + NEWF + MAJORPTYF + AGE), data = imai,
                  method = "cbps", estimand = "ATE")
```

## Create Plots of Distribution of Key Variables

```{r, echo = TRUE, size="tiny", fig.height=4}
bal.plot(W.out, var.name = "AGE")
```

## Create Plots of Distribution of Key Variables

```{r, echo = TRUE, size="tiny", fig.height=4}
bal.plot(W.out, var.name = "VOTE96.1F")
```